---
title: "p8105_hw3"
author: "qz2492"
date: "2022-10-11"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```


### Problem 2

#### Load, tidy, and wrangle the data.
```{r}

accel = read_csv('./data/accel_data.csv')
accel_clean = accel %>% 
  pivot_longer(
    cols = activity.1:activity.1440,
    names_to = "number",
    values_to = "counts",
    names_prefix = "activity."
  ) %>% 
  mutate(
    weekend = (day %in% c('Saturday', 'Sunday'))
  ) %>%
  relocate(week, day_id, weekend, day)

head(accel_clean, 7) %>%
  knitr::kable()
```

The `accel` dataset contains `r nrow(accel_clean)` observations and `r ncol(accel_clean)` variables, which are `week`, `day_id`, `day`, `weekend`, `number` and `counts`. All these variables are records of five weeks of accelerometer data collected on a 63 year-old male with BMI 25.

#### aggregate across minutes to create a total activity variable for each day

```{r}
accel_aggr = accel_clean %>%
  group_by(week, day) %>%
  summarise(activity_sum = sum(counts)) %>%
  pivot_wider(
    names_from = "day",
    values_from = "activity_sum"
  ) %>%
  relocate(week, Sunday, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday) 
accel_aggr
```

There is no significant trends in this 5-weeks activities record, but we can tell that on 4th and 5th Saturdays the activity frequency was reduced to 1 time per minute, perhaps it was because he removed his equipment.
#### Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week

```{r}
accel_clean %>%
  mutate(number = as.numeric(number)) %>%
  group_by(day, number) %>% 
  summarize(avg_counts = mean(counts)) %>%
  ggplot(aes(x = number, y = avg_counts, color = day)) +
  geom_smooth(alpha = 0.2) +
  theme(legend.position = "right") +
  scale_x_continuous(
    breaks = c(0:24) * 60,
    labels = c(0:24)
  ) +
  labs(
    title = "24-hour activity time courses for each day",
    x = "Hours for each day",
    y = "Average activity counts")
  

```
  
Based on this graph, we can tell that this 63 year-old male reaches his activity summit at about 9 o'clock on Friday night, and he is the second most active at about 10 o'clock on Sunday morning. Besides, his activity frequency is usually low during 11 p.m. to 4 a.m and the average activity times during the day is approximately 400.

### Problem 3
```{r}
data("ny_noaa")
```

The raw dataset `ny_noaa` contains `r nrow(ny_noaa)` observations with `r ncol(ny_noaa)` variables, which include `id`, `date`, `prcp`, `snow`, `snwd`, `tmax` and `tmin`. `id` means weather station ID, while `date` represents date of observation. `prcp` means precipitation (tenths of mm), `snow` means snowfall (mm) and `snwd` means snow depth (mm). `tmax` and `tmin` each represent the maximum temperature and minimum temperature(tenths of degrees C) for the day. 

#### Do some data cleaning
Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
```{r}
ny_tidy = ny_noaa %>%
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(
    tmax_c = as.double(tmax) / 10,
    tmin_c = as.double(tmin) / 10,
    prcp_c = as.double(prcp) / 10
  )

ny_tidy %>%
  group_by(snow) %>%
  summarise(n_obs = n()) %>%
  arrange(desc(n_obs)) %>%
  head(3)
```
The most commonly observed values for snowfall is 0, because for NYC, snow often occurs in winter.

#### Make a two-panel plot showing the average max temperature in January and in July in each station across years.
```{r}
ny_tidy %>% 
  filter(month == "01" | month == "07") %>%
  group_by(id, year, month) %>%
  summarize(avg_tmax = mean(tmax_c, rm.na = TRUE)) %>%
  ggplot(aes(x = as.integer(year), y = avg_tmax, color = id)) +
  geom_point(alpha = 0.6) +
  geom_line(alpha = 0.4) +
  scale_x_continuous(breaks = c(1980, 1990, 2000, 2010)) +
  theme(legend.position = "none") +
  labs(
    title = "The average max temperature in January and in July",
    x = "Years",
    y = "The average max temperature(Celsius)"
  ) +
  facet_grid(~month)
  
```

We can tell from the picture that the average max temperature in July is significantly higher than it in January. Besides, the average max temperature in every station of NYC seemed to have the same tendency of temperature changes, and this had nothing to do with month. The min temperature is in the range of -10 to 10(Celsius) and the max temperature is in the range of 20 to 35(Celsius). The lines in this picture are approximately parallel. But there are still some outliers like the point in January, 1982 and the point in July, 1988, which are extremely lower than the average.

#### Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
```{r}
# tmax vs tmin for the full dataset
tmax_vs_tmin = ny_tidy %>%
  ggplot(aes(x = tmax_c, y = tmin_c)) +
  geom_hex() +
  theme(legend.position = "right") +
  labs(
    title = "tmax vs tmin",
    x = "Max temperature(Celsius)",
    y = "Max temperature(Celsius)")
# make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year
snowfall = ny_tidy %>%
  filter(snow > 0 & snow < 100) %>%
  ggplot(aes(x = snow, y = year)) +
  geom_density_ridges(color = 'burlywood1') +
  scale_x_continuous(limit = c(0, 100)) +
  labs(
    title = "Snowfall distribution",
    x = "Snowfall(mm)",
    y = "Year"
  )
tmax_vs_tmin + snowfall
```

