---
title: "p8105_hw3"
author: "qz2492"
date: "2022-10-11"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```


### Problem 2

#### Load, tidy, and wrangle the data.
```{r}

accel = read_csv('./data/accel_data.csv')
accel_clean = accel %>% 
  pivot_longer(
    cols = activity.1:activity.1440,
    names_to = "number",
    values_to = "counts",
    names_prefix = "activity."
  ) %>% 
  mutate(
    weekend = (day %in% c('Saturday', 'Sunday'))
  ) %>%
  relocate(week, day_id, weekend, day)

head(accel_clean, 7) %>%
  knitr::kable()
```

The `accel` dataset contains `r nrow(accel_clean)` observations and `r ncol(accel_clean)` variables, which are `week`, `day_id`, `day`, `weekend`, `number` and `counts`. All these variables are records of five weeks of accelerometer data collected on a 63 year-old male with BMI 25.

#### aggregate across minutes to create a total activity variable for each day

```{r}
accel_aggr = accel_clean %>%
  group_by(week, day) %>%
  summarise(activity_sum = sum(counts)) %>%
  pivot_wider(
    names_from = "day",
    values_from = "activity_sum"
  ) %>%
  relocate(week, Sunday, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday) 
accel_aggr
```

There is no significant trends in this 5-weeks activities record, but we can tell that on 4th and 5th Saturdays the activity frequency was reduced to 1 time per minute, perhaps it was because he removed his equipment.
#### Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week

```{r}
accel_clean %>%
  mutate(number = as.numeric(number)) %>%
  group_by(day, number) %>% 
  summarize(avg_counts = mean(counts)) %>%
  ggplot(aes(x = number, y = avg_counts, color = day)) +
    geom_smooth(alpha = 0.2) +
    theme(legend.position = "right") +
    scale_x_continuous(
      breaks = c(0:24) * 60,
      labels = c(0:24)
    ) +
    labs(
      title = "24-hour activity time courses for each day",
      x = "Hours for each day",
      y = "Average activity counts")
  

```
  
Based on this graph, we can tell that this 63 year-old male reaches his activity summit at about 9 o'clock on Friday night, and he is secondly active at about 10 o'clock on Sunday morning. Besides, his activity time is usually low during 11 p.m. to 4 a.m.


